Bioverse: Potentially Observable Exoplanet Biosignature Patterns Under the UV Threshold Hypothesis for the Origin of Life

### 2025-02-05 src/tex/ms.tex
edited discusssion (and polished methods)

\rev{Then, we turn to a metric which measures the information gain of sampling: the Bayes factor \cite{Jeffreys1939}.
The posterior evidence for the null hypothesis $H_{\mathrm{null}}$ is given by the Bayes' theorem:
\begin{equation}
\label{eq:bayesHnull}
    P(H_{\mathrm{null}}|Y) = \frac{P(Y|H_{\mathrm{null}})P(H_{\mathrm{null}})}{P(Y)}
\end{equation}
and the posterior evidence for the NUV hypothesis:
\begin{equation}
    \label{eq:bayesHone}
        P(H_1|Y) = \frac{P(Y|H_1)P(H_1)}{P(Y)}~.
\end{equation}
Computing the distributions given by equations \ref{eq:bayesHnull} and \ref{eq:bayesHone} requires the distributions of likelihoods, given in equations \ref{eq:semian:likelihoodHnull} and \ref{eq:semian:likelihoodH1}, as well as the prior distributions for each hypothesis ($P(H_{\mathrm{null}})$ and $P(H_{1})$) and for the sample ($P(H_{1})$).
}
\rev{
Apart from the caveat that the likelihood for $H_{1}$ cannot be expressed analytically due to being conditioned by n_{\lambda} (equation \ref{eq:semian:likelihoodH1}), computing posterior probabilities is challenging due to the unknown priors.
In addition, we are not interested in constraining the posterior distributions for each hypothesis as we are in measuring the information gain of performing a certain observation, regardless of the priors that future observers may adopt.
The ratio of posterior probabilities (equations \ref{eq:bayesHone} and \ref{eq:bayesHnull}) is the product of the ratio of the prior probabilities and likelihoods of each hypotheses (equation \ref{eq:postratio}).
\begin{equation}
    \label{eq:postratio}
    \frac{P(H_1|X)}{P(H_{\mathrm{null}}|X)} = \frac{P(X|H_{\mathrm{null}})P(H_{\mathrm{null}})}{P(X|H_{\mathrm{null}})P(H_{\mathrm{null}})}~.
\end{equation}
By truncating the prior ratio out of equation \ref{eq:postratio} (and substituting likelihoods for their expressions; equations \ref{eq:semian:likelihoodH1}, \ref{eq:semian:likelihoodHnull}), one obtains the Bayes factor \cite{Jeffreys1939}:
}
%\begin{subequations}
\begin{equation}\label{eq:bayes_factor}
BF_{H_1,H_\mathrm{null}} = \frac{P(Y=k|H_1)}{P(Y=k|H_\mathrm{null})} = \frac{\binom{n_\lambda}{k}}{\binom{n}{k}}(1-f_\mathrm{life})^{n_{\lambda}-n},
\end{equation}
\rev{and conversely the Bayes ratio measuring evidence for $H_{\mathrm{null}}$ against $H_1$:}
\begin{equation}
    \label{eq:bayes_factor2}
BF_{H_\mathrm{null},H_1} = \frac{P(Y=k|H_\mathrm{null})}{P(Y=k|H_1)} = \frac{\binom{n}{k}}{\binom{n_\lambda}{k}}(1-f_\mathrm{life})^{n-n_{\lambda}}.
\end{equation}
%\end{subequations}
\rev{
By focusing on the Bayes ratio, we measure how much the observation of a sample $Y$ tips the scale of evidence for or against hypothesis relative to prior expectations, allowing us to focus on the information yield as a function of sample size.}

@@

>>>>

\rev{Then, we turn to a metric which measures the information gain of sampling: the Bayes factor \cite{Jeffreys1939}.
The posterior evidence for the null hypothesis $H_{\mathrm{null}}$ is given by the Bayes' theorem:
\begin{equation}
\label{eq:bayesHnull}
    P(H_{\mathrm{null}}|Y) = \frac{P(Y|H_{\mathrm{null}})P(H_{\mathrm{null}})}{P(Y)}
\end{equation}
and the posterior evidence for the NUV hypothesis:
\begin{equation}
    \label{eq:bayesHone}
        P(H_1|Y) = \frac{P(Y|H_1)P(H_1)}{P(Y)}~.
\end{equation}
Computing the distributions given by equations \ref{eq:bayesHnull} and \ref{eq:bayesHone} requires the distributions of likelihoods, given in equations \ref{eq:semian:likelihoodHnull} and \ref{eq:semian:likelihoodH1}, as well as the prior distributions for each hypothesis ($P(H_{\mathrm{null}})$ and $P(H_{1})$) and for the sample ($P(H_{1})$).
}
\rev{
Apart from the caveat that the likelihood for $H_{1}$ cannot be expressed analytically due to being conditioned by n_{\lambda} (equation \ref{eq:semian:likelihoodH1}), computing posterior probabilities is challenging due to the unknown priors.
In addition, we are not interested in constraining the posterior distributions for each hypothesis as we are in measuring the information gain of performing a certain observation, regardless of the priors that future observers may adopt.
Previous research aimed at prospectively quantifying the information yield of future exoplanet surveys focused on the Bayes factor, rather than on posterior probabilities \cite{affholder2025interior}.
The ratio of posterior probabilities (equations \ref{eq:bayesHone} and \ref{eq:bayesHnull}) is the product of the ratio of the prior probabilities and likelihoods of each hypotheses (equation \ref{eq:postratio}).
\begin{equation}
    \label{eq:postratio}
    \frac{P(H_1|Y)}{P(H_{\mathrm{null}}|Y)} = \frac{P(Y|H_{\mathrm{null}})P(H_{\mathrm{null}})}{P(Y|H_{\mathrm{null}})P(H_{\mathrm{null}})}~.
\end{equation}
By truncating the prior ratio out of equation \ref{eq:postratio} (and substituting likelihoods for their expressions; equations \ref{eq:semian:likelihoodH1}, \ref{eq:semian:likelihoodHnull}), one obtains the Bayes factor \cite{Jeffreys1939}:
}
%\begin{subequations}
\begin{equation}\label{eq:bayes_factor}
BF_{H_1,H_\mathrm{null}} = \frac{P(Y=k|H_1)}{P(Y=k|H_\mathrm{null})} = \frac{\binom{n_\lambda}{k}}{\binom{n}{k}}(1-f_\mathrm{life})^{n_{\lambda}-n},
\end{equation}
\rev{and conversely the Bayes ratio measuring evidence for $H_{\mathrm{null}}$ against $H_1$:}
\begin{equation}
    \label{eq:bayes_factor2}
BF_{H_\mathrm{null},H_1} = \frac{P(Y=k|H_\mathrm{null})}{P(Y=k|H_1)} = \frac{\binom{n}{k}}{\binom{n_\lambda}{k}}(1-f_\mathrm{life})^{n-n_{\lambda}}.
\end{equation}
%\end{subequations}
\rev{
By focusing on the Bayes ratio, we measure how much the observation of a sample $Y$ tips the scale of evidence for or against hypothesis relative to prior expectations, allowing us to focus on the information yield as a function of sample size.}

@@

\rev{
Our semi-analytical framework focuses on the Bayes ratio as a measure of the evidence that a future observation may bring relative to prior expectations.
This is useful in order to specifically consider the information yield of an observation, but also trades off the opportunity to estimate the posterior probability of a specific hypothesis with the possibility to narrow down the problem to two specific competing hypotheses.
Indeed, the posterior probability depends on the prior expectation for a sample $Y$ (see equation \ref{eq:bayesHone}), which may be approximated as the sum of likelihoods of the modeled hypotheses ($P(Y)=P(Y|H_1)P(H_1)+P(Y|H_{\mathrm{null}})P(H_{\mathrm{null}})$) but could in principle depend on any number of unknown processes.
Could there be factors other than the past UV flux that would determine the occurrence rate of biosignatures in terrestrial exoplanets?
Surely, there is a vast array of possible mechanisms that could be considered which could shape the prior probability of an observed sample of biosignatures $Y$.
Thus, while the Bayes factor is an essential part of the inference process, and the one relevant to prospective analyses of future evidence yields, future attempts at inference may aim for estimating posterior probabilities rather than evidence ratios.
Constraining the prior distribution for the probability that life requires exposure to a certain NUV threshold to emerge is beyond the scope of this study, but remains a question that future work will need to address in order to evaluate fully this hypothesis in the prospect of future observation of biosignatures.}
